Wikipedia Text Mining and Entropy Analysis

### 2019-01-30 cogsci/position-info.Rmd
editing cogsci paper in progress

```{r, libraries}
library(png)
library(grid)
library(tidyverse)
library(xtable)
library(tidytext)
library(tidyboot)
library(entropy)
```

@@

```{r our_replication, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}

img <- png::readPNG("figs/Prov_corpus_result.png")
grid::grid.raster(img)
```

@@

@yu2016 challenges the UID hypothesis through examining an entropy measure across sentence positions within the text portion of the English-language British National Corpus (BNC) @clear1993british. They partition the corpus by sentence length in number of words. For each word position $X$ of sentences of length $k$, they define $w$ as a unique word occurring in position $X$. They define $p(w)$ as the number of times word $w$ occurs in position $X$ divided by the number of total words that occur in position $X$ i.e. the number of sentences of length $k$. Then $p(w)$ is the probability of obtaining word $w$ by choosing a word at random in position $X$ in sentences of length $k$.

@@

With this measure, @yu2016 compute the unigram entropy at each position of sentences of each length within the corpus. The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances. @genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus. UID applies this uniformity of entropy rate in sentences to all levels of speech, and so the @yu2016, which examines text at the word level, should find an affine function at the word level.

@@

We first used the same source for data as @levy2007, the Switchboard corpus of American telephone conversations @godfrey1992.

```{r switchboard, include=FALSE}
library(feather)

sb_tokens <- read_feather("../switchboard/switchboard.feather") %>%
  mutate(length = str_count(value, pattern = " +") + 1) %>% 
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, value, token = stringr::str_split, pattern = " +") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

unigram_entropy <- function(sen_length, tokens) {
  tokens %>%
    filter(length == sen_length) %>%
    group_by(word_order, word) %>%
    summarise(n = n()) %>%
    tidyboot(summary_function = function(x) x %>% 
               summarise(entropy = entropy(n, unit = "log2")),
             statistics_functions = function(x) x %>%
             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
    mutate(length = sen_length)
}

entropies <- map(c(5, 7, 9), ~unigram_entropy(.x, sb_tokens)) %>%
  bind_rows()

ggplot(entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  ggtitle("Representative Switchboard Entropy Curves Faceted on Utterance Length") + 
  xlab("Word Order within Utterance") +
  ylab("Positional Entropy") +
  scale_x_discrete(limits = "1":"9") + 
  geom_pointrange() +
  geom_smooth(se = F)
```

@@

We used the Providence English corpus from CHILDES. The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home. For speech data, which for the corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora.We divided the corpus by speaker into child and non-child categories. We further divided the corpus by utterance length, so that all sentences of length $k$ (e.g. $6$) were grouped together. Finally, within each utterance length, we computed the unigram entropy measure for each position. Computation of positional entropy was identical to the Switchboard method after dividing for utterance length.

```{r providence_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}
img <- png::readPNG("figs/Prov_corpus_result.png")
grid::grid.raster(img)
```

@@

@@

## Results & Analysis
The adult and child entropy curves track one another almost identically. This is surprising because UID would predict that the young children in the corpora who are not fully developed speakers would have more noise in their distribution of information rates. This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees. We also observed a robustness across corpora for the same languages, and a robustness across utterance lengths within the same corpus's entropy curve. This shows that the concept of an "entropy curve" for a specific language is well-founded when considering speech data.

@@

The UID hypothesis also applies to written communication: we expect people to communicate information at a uniform rate through writing as well. We use Wikipedia as a source for written data, which provides two advantages. One, the quantity of data in Wikipedia is large for each language and two, there are hundreds of languages with Wikipedias. This allows us to perform the entropy analysis on a much greater scale and to directly compare the results of the entropy analysis on each language to one another, and ultimately to predict what typological features of a language help determine its entropy curve, if any. We will describe our method of harvesting and distilling text data from each Wikipedia as well as how we compare the entropy curves from each language to one another.

@@

Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extract text corpora from Wikipedia by downloading a stored collection of Wikipedia entries in each langauge and randomly selecting several thousand articles from each Wikipedia language. Each language corpus was cleaned and limited to sentences between $6$ and $50$ words. Similar to our process for the Switchboard and CHILDES spoken corpora, we divided each corpus by sentence length, and then computed the unigram entropy measure on each word position within each sentence length. We wanted to directly compare and classify the unigram entropy curves of the languages from Wikipedia. We computed three slope treatments of each curve. In the *absolute* treatment, with sentence length denoted as $n$, we computed the slope between positions $1$ and $2$, positions $2$ and $3$, positions $3$ and $n-2$, positions $n-2$ and $n-1$ and positions $n-1$ and $n$. For the short utterances appearing the CHILDES speech corpora we examined, these appeared to be important junctions in the distributions, with a seeming plateau in the middle of the unigram entropy curve for each of the language corpora we examined in CHILDES.

However, because the portion of sentences of length greater than $10$ in the Wikipedia corpora were significantly larger than the CHILDES corpora, then we also computed relative slope treatments. In the *relative 5* treatment, we computed the slopes between $0\%$ and $20\%$, $20\%$ and $40\%$, $40\%$ and $60\%$, $60\%$ and $80\%$ and $80\%$ and $100\%$ of the relative word positions in each sentence. When one of these percentages was not a whole number, then the closest whole number position was used instead for slope calculation. In the *relative 10* treatment, we computed the slopes between every $10\%$ of the relative word positions in each sentence. Each comparative slope within each treatment was averaged together between different sentence lengths, for example in the *relative 5* treatment then all of the $0\%$ to $20\%$ slopes were averaged together. This created three treatments for the entropy curve for each langauge in the Wikipedia database.

@@

In the *absolute* and *relative 5* treatments, each language is embedded in $5$-dimensional space. In the *relative 10* treatment, each language is embedded in $10$-dimensional space. This allows for direct comparison between languages on Wikipedia using cosine similarity and moreover for unsupervised clustering analysis to compare the results of our position-wise analysis on datasets from Wikipedia to known typological features within the languages. To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we use the linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013).

[Ideally do everything, but missing values in WALS, and how to compare every combination of features] An ideal approach using all of the corpora we can download from Wikipedia and all $144$ features from the WALS database. Cluster the languages used in the Wikipedia analysis on the basis of WALS features and then directly compare the WALS clustering results with the results of the hierarchical clustering on the Wikipedia slope data for each of the different treatments using clustering similarity evaluation methods such as the Rand index (Rand, 1971). However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis. This is a computationally intractable operation. Additionallly, deciding how many clusters to use in an unsupervised clustering analysis is an unsolved problem in machine learning.

We instead check the effects of individual features on the embeddings of languages in the different treatments. We computed pairwise cosine similarity between each pair of language vectors within a treatment. For a subset of WALS features which had values entered for most of the languages we obtained from Wikipedia, we used a generalized linear model to see whether the cosine similarity between languages mattered in predicting if the languages shared the same value for a WALS feature. For eight features and $45$ languages, we found this to be true. The table below shows the results for the generalized linear model we computed using WALS features and the absolute treatment.

@@

@@

# Discussion 
In this paper, we have extended and applied a model from @yu2016 and derived from @genzel2002 for distinguishing entropy at each word position within sentences. We have argued that the outcomes of this model are derived from the information distribution of a language. As this model varies cross-linguistically in terms of

Follow-up with WALS features using missing-value imputation. However here are some of the problems with that.

Our work complements the approach of studies such as @aylett2004, where languages are characterized by a rate of semantic information per syllable. This body of work attests to languages possessing different rates of information transfer based on typological features as well.

We also propose follow-up within linguistic processing using evidence from cross-linguistic eye-tracking. Research has indicated the effects of surprisal on fixation duration during eye-tracking studies [@demberg2008; @boston2008; @smith2013]; higher surprisal of a word is on average predictive of higher fixation duration. The *wrap-up effect* states that when a person reads written text, he or she will process sentence-final words more slowly on average then sentence-medial or sentence-initial words, due to integrating information from the entire sentence to form a final understanding of the sentence's meaning [@stowe2018; @kuperman2010]. The wrap-up effect is drawn from evidence in languages with a large final increase in their entropy curve from our study, so perhaps the wrap-up effect derives from the sentence-final increase in entropy curves observed in English and other Germanic languages.

>>>>

```{r, libraries}
library(png)
library(grid)
library(tidyverse)
library(xtable)
library(tidytext)
library(tidyboot)
library(entropy)
library(here)
library(feather)
library(magrittr)
```

@@

@@

Here we will describe our adaptation of the by-word entropy method in @yu2016, which uses the text portion of the British National Corpus @clear1993british. We replicate their results at the end of this subsection. Given a text or speech corpus divided into individual utterances, we partition the corpus by utterance length in number of words. For each word position $X$ of utterances of length $k$, we define $w$ as a unique word occurring in position $X$. We further define $p(w)$ as the number of times word $w$ occurs in position $X$, divided by the number of total words that occur in position $X$ i.e. the number of sentences of length $k$. This creates a probability distribution over the words occurring in position $X$, and computing the Shannon entropy $H(X)$ @shannon1948 of this probability distribution gives the positional entropy of position $X$ in utterances of length $k$.

@@

Our replication of the @yu2016 results

```{r read_and_plot_bnc}
bnc_entropies <- read_csv(here("data/bnc_entropies.csv"))

ggplot(bnc_entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  ggtitle("Representative BNC Entropy Curves Faceted on Utterance Length") + 
  xlab("Word Order within Utterance") +
  ylab("Positional Entropy") +
  geom_pointrange() +
  geom_line()
```

With this measure, we compute the unigram entropy at each position of sentences of each length within the corpus. The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances. @genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus. UID applies this uniformity of entropy rate in sentences to all levels of speech, and so our method obtained from @yu2016, which examines text at the word level, should find an affine function at the word level.

@@

We first used the same source for data as @levy2007, the Switchboard corpus of American telephone conversations @godfrey1992. The Switchboard corpus contains telephone conversations of American English between strangers, with individual utterances ranging from one word to several dozen words. We divided the Switchboard corpus by utterance length and computed the positional entropy for each position within each sentence length, then plotted the result as an entropy curve.

Our plots for three representative utterance lengths are below.

```{r read_and_plot_switchboard}
switchboard_entropies <- read_csv(here("data/switchboard_entropies.csv")) %>%
  filter(length %in% c(5, 7, 9))

ggplot(switchboard_entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  ggtitle("Representative Switchboard Entropy Curves Faceted on Utterance Length") + 
  xlab("Word Order within Utterance") +
  ylab("Positional Entropy") +
  scale_x_discrete(limits = "1":"9") + 
  geom_pointrange() +
  geom_smooth(se = F)
```

@@

After replicating the @yu2016 results using the spoken conversation corpus from @levy2007, we want to determine whether the entropy curves will remain robust over child-adult conversational corpora.

The corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora. For our first corpus, we used the Providence English corpus from CHILDES @macwhinney2014. The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home. We divided the corpus by speaker into child and non-child categories. We further divided the corpus by utterance length, so that all sentences of length $k$ (e.g. $6$) were grouped together. Finally, within each utterance length, we computed the unigram entropy measure for each position. Computation of positional entropy was identical to our general method after dividing for utterance length.

@@

## Results and Analysis
```{r providence_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}
img <- png::readPNG("figs/Prov_corpus_result.png")
grid::grid.raster(img)
```

@@

The adult and child entropy curves track one another almost identically. This is surprising because UID would predict that the young children in the corpora who are not fully developed speakers would have more noise in their distribution of information rates. This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees. We also observed a robustness across corpora for the same languages, and a robustness across utterance lengths within the same corpus's entropy curve. This shows that the concept of an "entropy curve" for a specific language is well-founded when considering speech data.

@@

The UID hypothesis also applies to written communication: we expect people to communicate information at a uniform rate through writing as well. We use Wikipedia as a source for written data, which provides two advantages. One, the quantity of data in Wikipedia is very large for each language, at least several megabytes, and two, there are hundreds of languages with their own Wikipedia corpora. We can therefore run our entropy analysis on a much greater scale than using small spoken CHILDES corpora and directly compare the results of the entropy analysis on each language to one another. We expect to find linguistic features that determine the shape of a language's entropy curve. We treated sentences in the Wikipedia text corpora as equivalent to utterances in the Switchboard and CHILDES corpora.

@@

Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extract text corpora from Wikipedia by downloading a stored collection of Wikipedia entries in each langauge and randomly selecting several thousand articles from each Wikipedia language. Each language corpus was cleaned and limited to sentences between six and $50$ words. Similar to our process for the Switchboard and CHILDES spoken corpora, we divided each corpus by sentence length, and then computed our positional entropy measure on each word position within each sentence length.

We computed two slope treatments of each curve. In the *absolute* treatment, with sentence length denoted as $k$, we computed the slope between positions $1$ and $2$, positions $2$ and $3$, positions $3$ and $k-2$, positions $k-2$ and $k-1$ and positions $k-1$ and $k$. For the short utterances appearing the CHILDES speech corpora, the slopes between the first and second slopes on either end of the distributions appeared to be more characteristic of the distribution as a whole, with a plateau in the middle of the entropy curve for each of the language corpora we examined in CHILDES.

However, sentences of length greater than ten in the Wikipedia corpora were significantly more common than in the CHILDES. Sentences on Wikipedia are commonly , then we also computed relative slope treatments. In the *relative 5* treatment, we computed the slopes between $0\%$ and $20\%$, $20\%$ and $40\%$, $40\%$ and $60\%$, $60\%$ and $80\%$ and $80\%$ and $100\%$ of the relative word positions in each sentence. When one of these percentages was not a whole number, then the closest whole number position was used instead for slope calculation. Each comparative slope within each treatment was averaged together between different sentence lengths, for example in the relative $5$ treatment then all of the $0\%$ to $20\%$ slopes were averaged together. This created three treatments for the entropy curve for each langauge in the Wikipedia database.

In both the absolute and relative $5$ treatments, each language is embedded in $5$-dimensional space. This allows for direct comparison between languages on Wikipedia using cosine similarity and moreover for unsupervised clustering analysis to compare the results of our position-wise analysis on datasets from Wikipedia to known typological features within the languages. To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we use the linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013).

We check the effects of individual features on the embeddings of languages in the different treatments. We computed pairwise cosine similarity between each pair of language vectors within a treatment. For a subset of WALS features which had values entered for most of the languages we obtained from Wikipedia, we used a generalized linear model to see whether the cosine similarity between languages mattered in predicting if the languages shared the same value for a WALS feature.

@@

For eight features and $45$ languages, we found this to be true. The table below shows the results for the generalized linear model we computed using WALS features and the absolute treatment.

@@

We also used the linear model approach to evaluate the effects of cosine distance in the relative $5$ treatment on the values of the same WALS features and found similar outcomes.

@@

# Discussion 
In this paper, we have extended and applied a model from @yu2016 and derived from @genzel2002 for distinguishing entropy at each word position within sentences. We have argued that the outcomes of this model are derived from the distribution of information transmission in human communication. Language is not a uniform noisy channel of communication, but each language individually represents a different noisy channel of communication characterized in part by typological features.

Our work complements the approach of studies such as @aylett2004, where languages are characterized by a single number, representing the rate of semantic information transfer per syllable. This body of work attests to languages possessing different rates of information transfer based on typological features as well. Within the information transfer rate literature, the rate is an overall feature of language that does not vary corpus to corpus. Similarly, in our work we have obtained a characteristic entropy distribution for each language. Both information rate and entropy distribution are characterized by typological features. This indicates that the structure and rate of how people communicate information varies cross-linguistically based on typological features.

An ideal approach would use all of the corpora we can download from Wikipedia and all $144$ features from the WALS database. Cluster the languages used in the Wikipedia analysis on the basis of WALS features and then directly compare the WALS clustering results with the results of the hierarchical clustering on the Wikipedia slope data for each of the different treatments using clustering similarity evaluation methods such as the Rand index (Rand, 1971). However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis. This is a computationally intractable operation. Additionallly, deciding how many clusters to use in an unsupervised clustering analysis is an unsolved problem in machine learning.

On a more mechanical level, we expect that our entropy model will have implications for how long people take to read individual words at different positions of a sentence. Eye-tracking research has indicated the effects of surprisal on fixation duration during eye-tracking studies [@demberg2008; @boston2008; @smith2013]; higher surprisal of a word is on average predictive of higher fixation duration. The *wrap-up effect* states that when a person reads written text, he or she will process sentence-final words more slowly on average then sentence-medial or sentence-initial words, due to integrating information from the entire sentence to form a final understanding of the sentence's meaning [@stowe2018; @kuperman2010]. The wrap-up effect is drawn from evidence in languages with a large final increase in their entropy curve from our study, so perhaps the wrap-up effect derives from the sentence-final increase in entropy curves observed in English and other Germanic languages.

### 2019-01-30 cogsci/position-info.tex
editing cogsci paper in progress

{\centering \includegraphics{figs/our_replication-1}

\caption[Providence corpus unigram entropy]{Providence corpus unigram entropy}\label{fig:our_replication}
\end{figure*}
\end{CodeChunk}

@@

The UID hypothesis predicts that information transmission rate will tend
towards uniformity in both speech and text. We begin by examining speech
to determine if we obtain the entropy curve shape we expect for spoken
communication. We begin with the same source as R. P. Levy \& Jaeger
(2007) and Jaeger (2010), the Switchboard corpus of adult telephone
conversations. We also examine child-adult conversations in the CHILDES
TalkBank (Brown, 1973; MacWhinney, 2014) corpora database of spoken
adult-child conversations in multiple languages. Children are not fully
developed speakers, so we want to compare the entropy curve we obtain by
computing over the utterances in the CHILDES corpora to the utterances
in the adult-adult Switchboard corpus.

@@

Yu et al. (2016) challenges the UID hypothesis through examining an
entropy measure across sentence positions within the text portion of the
English-language British National Corpus (BNC) Clear (1993). They
partition the corpus by sentence length in number of words. For each
word position \(X\) of sentences of length \(k\), they define \(w\) as a
unique word occurring in position \(X\). They define \(p(w)\) as the
number of times word \(w\) occurs in position \(X\) divided by the
number of total words that occur in position \(X\) i.e.~the number of
sentences of length \(k\). Then \(p(w)\) is the probability of obtaining
word \(w\) by choosing a word at random in position \(X\) in sentences
of length \(k\).

\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\] With this measure, Yu et
al. (2016) compute the unigram entropy at each position of sentences of
each length within the corpus. The result of this method can be plotted
for each utterance length as an \emph{entropy curve}, which can be
visually compared across utterance length to observe the how the unigram
entropy changes across absolute positions in each of the utterances.
Genzel \& Charniak (2002) similarly examine a unigram entropy measure on
sentences, and found that entropy at the sentence level increases
linearly with sentence index within a corpus. UID applies this
uniformity of entropy rate in sentences to all levels of speech, and so
the Yu et al. (2016), which examines text at the word level, should find
an affine function at the word level.

@@

We used the Providence English corpus from CHILDES. The Providence
corpus recorded interactions between children between 1 and 3 years old
and their parents in the home. We divided the corpus by speaker into
child and non-child categories. We further divided the corpus by
utterance length, so that all sentences of length \(k\) (e.g. \(6\))
were grouped together. Finally, within each utterance length, we
computed the unigram entropy measure for each position.

The entropy curves capture individual variation across positions in
utterances of the same length. This allows us to directly observe and
judge the amount of variation in words that appear in an individual
position of a sentence. For speech data, which for the corpora in the
CHILDES database consists of short and often disconnected utterances
across hours of recordings, the unigram entropy measure is unaffected by
context or lack thereof in utterances by the adults and children in the
corpora. We can directly compare any two positions within utterances to
determine the amount of uncertainty, and therefore information, on
average contained by words within that position of utterances. We are
applying the same approach as in Genzel \& Charniak (2002), but within
sentences instead of across sentences.

@@

\subsection{Results \& Analysis}\label{results-analysis}

@@

However, because the portion of sentences of length greater than \(10\)
in the Wikipedia corpora were significantly larger than the CHILDES
corpora, then we also computed relative slope treatments. In the
\emph{relative 5} treatment, we computed the slopes between \(0\%\) and
\(20\%\), \(20\%\) and \(40\%\), \(40\%\) and \(60\%\), \(60\%\) and
\(80\%\) and \(80\%\) and \(100\%\) of the relative word positions in
each sentence. When one of these percentages was not a whole number,
then the closest whole number position was used instead for slope
calculation. In the \emph{relative 10} treatment, we computed the slopes
between every \(10\%\) of the relative word positions in each sentence.
Each comparative slope within each treatment was averaged together
between different sentence lengths, for example in the \emph{relative 5}
treatment then all of the \(0\%\) to \(20\%\) slopes were averaged
together. This created three treatments for the entropy curve for each
langauge in the Wikipedia database.

In the \emph{absolute} and \emph{relative 5} treatments, each language
is embedded in \(5\)-dimensional space. In the \emph{relative 10}
treatment, each language is embedded in \(10\)-dimensional space. This
allows for direct comparison between languages on Wikipedia using cosine
similarity and moreover for unsupervised clustering analysis to compare
the results of our position-wise analysis on datasets from Wikipedia to
known typological features within the languages. To determine which
phonological, morphological and syntactic features affected the
embedding of a language in the Wikipedia dataset, we use the linguistic
features in World Atlas of Language Structures (Dryer \& Haspelmath,
2013).

{[}Ideally do everything, but missing values in WALS, and how to compare
every combination of features{]} An ideal approach using all of the
corpora we can download from Wikipedia and all \(144\) features from the
WALS database. Cluster the languages used in the Wikipedia analysis on
the basis of WALS features and then directly compare the WALS clustering
results with the results of the hierarchical clustering on the Wikipedia
slope data for each of the different treatments using clustering
similarity evaluation methods such as the Rand index (Rand, 1971).
However, the problem then arises of which combination of WALS features
and how many features to include in the clustering analysis. This is a
computationally intractable operation. Additionallly, deciding how many
clusters to use in an unsupervised clustering analysis is an unsolved
problem in machine learning.

We instead check the effects of individual features on the embeddings of
languages in the different treatments. We computed pairwise cosine
similarity between each pair of language vectors within a treatment. For
a subset of WALS features which had values entered for most of the
languages we obtained from Wikipedia, we used a generalized linear model
to see whether the cosine similarity between languages mattered in
predicting if the languages shared the same value for a WALS feature.
For eight features and \(45\) languages, we found this to be true. The
table below shows the results for the generalized linear model we
computed using WALS features and the absolute treatment.

@@

>>>>

@@

The UID hypothesis predicts that information transmission rate will tend
towards uniformity in both speech and text. We begin by examining speech
to determine if we obtain the entropy curve shape we expect for spoken
communication. We begin with the same source as R. P. Levy \& Jaeger
(2007) and Jaeger (2010), the Switchboard corpus of adult telephone
conversations. We also examine child-adult conversations in the CHILDES
TalkBank (Brown, 1973; MacWhinney, 2014) corpora database of spoken
adult-child conversations in multiple languages. Children are not fully
developed speakers, so we want to compare the entropy curve we obtain by
computing over the utterances in the CHILDES corpora to the utterances
in the adult-adult Switchboard corpus. Finally, we extend our
cross-linguistic findings on a small number of languages from the
CHILDES experiment to a direct comparison of several dozen languages in
Wikipedia.

\subsection{General methods}\label{general-methods}

Here we will describe our adaptation of the by-word entropy method in Yu
et al. (2016), which uses the text portion of the British National
Corpus Clear (1993). We replicate their results at the end of this
subsection. Given a text or speech corpus divided into individual
utterances, we partition the corpus by utterance length in number of
words. For each word position \(X\) of utterances of length \(k\), we
define \(w\) as a unique word occurring in position \(X\). We further
define \(p(w)\) as the number of times word \(w\) occurs in position
\(X\), divided by the number of total words that occur in position \(X\)
i.e.~the number of sentences of length \(k\). This creates a probability
distribution over the words occurring in position \(X\), and computing
the Shannon entropy \(H(X)\) Shannon (1948) of this probability
distribution gives the positional entropy of position \(X\) in
utterances of length \(k\).

\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\]

Our replication of the Yu et al. (2016) results

\begin{CodeChunk}

\includegraphics{figs/read_and_plot_bnc-1} \end{CodeChunk}

With this measure, we compute the unigram entropy at each position of
sentences of each length within the corpus. The result of this method
can be plotted for each utterance length as an \emph{entropy curve},
which can be visually compared across utterance length to observe the
how the unigram entropy changes across absolute positions in each of the
utterances. Genzel \& Charniak (2002) similarly examine a unigram
entropy measure on sentences, and found that entropy at the sentence
level increases linearly with sentence index within a corpus. UID
applies this uniformity of entropy rate in sentences to all levels of
speech, and so our method obtained from Yu et al. (2016), which examines
text at the word level, should find an affine function at the word
level.

The entropy curves capture individual variation across positions in
utterances of the same length. This allows us to directly observe and
judge the amount of variation in words that appear in an individual
position of a sentence. We can directly compare any two positions within
utterances to determine the amount of uncertainty, and therefore
information, on average contained by words within that position of
utterances. We are applying the same approach as in Genzel \& Charniak
(2002), but within sentences instead of across sentences.

@@

We first used the same source for data as R. P. Levy \& Jaeger (2007),
the Switchboard corpus of American telephone conversations Godfrey,
Holliman, \& McDaniel (1992). The Switchboard corpus contains telephone
conversations of American English between strangers, with individual
utterances ranging from one word to several dozen words. We divided the
Switchboard corpus by utterance length and

\begin{CodeChunk}

\includegraphics{figs/read_and_plot_switchboard-1} \end{CodeChunk}

@@

After replicating the Yu et al. (2016) results using the spoken
conversation corpus from R. P. Levy \& Jaeger (2007), we want to
determine whether the entropy curves will remain robust over child-adult
conversational corpora.

The corpora in the CHILDES database consists of short and often
disconnected utterances across hours of recordings, the unigram entropy
measure is unaffected by context or lack thereof in utterances by the
adults and children in the corpora. For our first corpus, we used the
Providence English corpus from CHILDES MacWhinney (2014). The Providence
corpus recorded interactions between children between 1 and 3 years old
and their parents in the home. We divided the corpus by speaker into
child and non-child categories. We further divided the corpus by
utterance length, so that all sentences of length \(k\) (e.g. \(6\))
were grouped together. Finally, within each utterance length, we
computed the unigram entropy measure for each position. Computation of
positional entropy was identical to our general method after dividing
for utterance length.

\subsection{Results and Analysis}\label{results-and-analysis-1}

@@

@@

However, because the portion of sentences of length greater than \(10\)
in the Wikipedia corpora were significantly larger than the CHILDES
corpora, then we also computed relative slope treatments. In the
\emph{relative 5} treatment, we computed the slopes between \(0\%\) and
\(20\%\), \(20\%\) and \(40\%\), \(40\%\) and \(60\%\), \(60\%\) and
\(80\%\) and \(80\%\) and \(100\%\) of the relative word positions in
each sentence. When one of these percentages was not a whole number,
then the closest whole number position was used instead for slope
calculation. Each comparative slope within each treatment was averaged
together between different sentence lengths, for example in the
\emph{relative 5} treatment then all of the \(0\%\) to \(20\%\) slopes
were averaged together. This created three treatments for the entropy
curve for each langauge in the Wikipedia database.

In both the \emph{absolute} and \emph{relative 5} treatments, each
language is embedded in \(5\)-dimensional space. This allows for direct
comparison between languages on Wikipedia using cosine similarity and
moreover for unsupervised clustering analysis to compare the results of
our position-wise analysis on datasets from Wikipedia to known
typological features within the languages. To determine which
phonological, morphological and syntactic features affected the
embedding of a language in the Wikipedia dataset, we use the linguistic
features in World Atlas of Language Structures (Dryer \& Haspelmath,
2013).

An ideal approach would use all of the corpora we can download from
Wikipedia and all \(144\) features from the WALS database. Cluster the
languages used in the Wikipedia analysis on the basis of WALS features
and then directly compare the WALS clustering results with the results
of the hierarchical clustering on the Wikipedia slope data for each of
the different treatments using clustering similarity evaluation methods
such as the Rand index (Rand, 1971). However, the problem then arises of
which combination of WALS features and how many features to include in
the clustering analysis. This is a computationally intractable
operation. Additionallly, deciding how many clusters to use in an
unsupervised clustering analysis is an unsolved problem in machine
learning.

We instead check the effects of individual features on the embeddings of
languages in the different treatments. We computed pairwise cosine
similarity between each pair of language vectors within a treatment. For
a subset of WALS features which had values entered for most of the
languages we obtained from Wikipedia, we used a generalized linear model
to see whether the cosine similarity between languages mattered in
predicting if the languages shared the same value for a WALS feature.

\subsection{Results and Analysis}\label{results-and-analysis-2}

For eight features and \(45\) languages, we found this to be true. The
table below shows the results for the generalized linear model we
computed using WALS features and the absolute treatment.

@@

\hypertarget{ref-godfrey1992}{}
Godfrey, J. J., Holliman, E. C., \& McDaniel, J. (1992). SWITCHBOARD:
Telephone speech corpus for research and development. In
\emph{Acoustics, speech, and signal processing, 1992. icassp-92., 1992
ieee international conference on} (Vol. 1, pp. 517--520). IEEE.

### 2019-01-30 wiki_entropy/wiki_clustering.Rmd
working on Cogsci submission: bibtex drafted, clustering table here
editing cogsci paper in progress

```{r}
wiki_gs <- gs_title("Wikipedia_absolute")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

@@

cosines <- wkpd %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, cosine, -language1)

@@

```{r evaluate_clusters}
wals_clusters <- wals_clust["cluster"] %>% unlist(use.names= FALSE)
```

>>>>

```{r}
wiki_gs <- gs_title("Wikipedia_relative5")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

@@

cosines <- wkpd %>%
  column_to_rownames("Language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, cosine, -language1)

@@
